Using SMT engine to generate Symbolic Automata\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2017-02-02 2017-Xudong-SMT/OpenPNets-SMT.tex
premier jet...

%\usepackage{amsfonts}
%\usepackage{fancyhdr}
%\usepackage{cite}
%\usepackage{ifthen}
%\usepackage{amssymb}
%\usepackage{fancyhdr}
%\usepackage{pifont}
\usepackage{stmaryrd}
\usepackage{mathtools,mathpartir}
\usepackage{proof}
%\usepackage{setspace}
%\usepackage{indentfirst}
\usepackage{amsmath,amssymb,amscd,mathrsfs}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}
\usepackage{epsfig,color,subfigure,enumitem}
\newcommand{\TODO}[1]{\textcolor{red}{\textbf{[TODO:#1]}}}
\newcommand{\NOTE}[1]{\textcolor{blue}{\textbf{[NOTE:#1]}}}
\newcommand{\ERIC}[1]{\textcolor{blue}{#1}}
\newcommand{\LUDO}[1]{\textcolor{green}{#1}}
\newcommand{\coloncolon}{{:\hspace{-.2ex}:}}
\makeatletter
\newcommand{\raisemath}[1]{\mathpalette{\raisem@th{#1}}}
\newcommand{\raisem@th}[3]{\raisebox{#1}{$#2#3$}}
\makeatother

@@

\title{More equivalences for Open pNets\thanks{This work was partially 
funded by the Associated Team FM4CPS
  between INRIA and ECNU, Shanghai}}
\author{ Ludovic Henrio\inst{1}  \ \ \  Eric Madelaine\inst{1,2} \ \ \ Min Zhang\inst{3}}
\institute{Univ. of Nice Sophia Antipolis, CNRS, UMR 7271, 06900 Sophia Antipolis, France
	\and INRIA Sophia Antipolis M\'edit\'erann\'ee, BP 93, 06902 Sophia Antipolis, France
\and Shanghai Key Laboratory of Trustworthy Computing, ECNU, China}
\date{}                                           % Activate to display a given date or no date

@@

\begin{abstract}
  Strong bisimulation allows to compare open systems described using the pNets model.
  In practice, as happens in process algebras, it is too strong, and
  we need to define some coarser relations, taking into account
  invisible or internal moves.

  Open pNets have a notion of \emph{synchronised actions}
  generalizing the usual internal actions (e.g. $\tau$ of CCS, or $i$
  in Lotos).

  We use this notion to define an equivalence relation similar to
  the classical \emph{weak bisimulation}, and study its properties...

  refinement preorders ?

  Algorithms ?

  Which significant (small) use-case ?

@@

\ERIC{imported from FORTE'16, but many of these works also had weak
  version... we need to have another look}

@@

\paragraph{Contribution}

@@

@@

% \TODO{Check whether we really use all these notations}

@@

The semantics of open pNets will be defined  as an open automaton. An open
automaton is an automaton where each transition composes transitions of several LTSs with
action of some holes, the transition occurs if some predicates hold, and can involve a 
set of state modifications.
%\TODO{adopt a uniform notation for open transitions, almost each instance has a 
%different 
%notation! I suggest p,l,pr,po using \{\} for p,l,po as they are sets}
\begin{definition}[Open transitions]
	\label{def:OpenTransitions}
	An \emph{open transition} over a set  $(S_i,s_{0 i}, \rightarrow_i)^{i\in
	I}$ of LTSs, a
	set $J$ of holes with sorts $Sort_j^{j\in J}$, and a set of states $\mathcal{S}$ is a structure of the form:	
	\begin{mathpar}
	\inferrule*[myfraction=\reddottedrule]
	{\{s_i~{\xrightarrow{a_i}}_i ~s_i^{\prime}\}^{i\in I},
		\{\xrightarrow{b_j}_j\}^{j\in J}, \Pred, \Post}
	{s \xrightarrow {v}s'}
	\end{mathpar}
	Where $s, s'\in\mathcal{S}$ and for all
        $i\in I$, $s_i{\xrightarrow{a_i}}_i s_i^{\prime}$ is a transition of the
	LTS $(S_i,s_{0 i}, \rightarrow_i)$, and $\xrightarrow{b_j}_j$
        is a transition of the hole $j$, for any action $b_j$ in the
        sort $Sort_j$. \Pred\ is a predicate
	over the different variables of the
	terms, labels, and states $s_i$, $b_j$, $s$, $v$. \Post\ is a set of equations that 
	hold \emph{after the open transition}, they are represented as a substitution of the 
	form $\{x_k\gets e_k\}^{k\in K}$ 
	where $x_k$ are variables of $s'$, $s'_i$, and $e_k$ are expressions over the other 
	variables of the open transition.
\end{definition}

@@

	%% \TODO{ We should be careful here: after (re) reading "Huimin
	%% 	Lin, 'Symbolic Transition Systems with Assignements', Concur'96" I
	%% 	think handling assignments is not trivial, even for comparisons of pLTSs. }

@@

	\medskip
%        \TODO{may be explain how $\Pred(SV,a_i^{i\in I_k},b_j^{j\in
%            J_k},v)$ is built ? You mean more than what is written on previous page????}
	%%    \TODO{I have tentatively added the sort constraint on hole actions, that was
	%%not included in the first version... I'm unsure whether this is the best place to
	%%include it, because it may change the decidability conditions on predicates}

@@

\subsection{Computing and using open automata}
In this section we present a simple algorithm to construct the open
automaton representing the behaviour of an open pNet, and we prove that
under reasonable conditions this automaton is finite.

\TODO{Eric: refine the algorithm description}

@@

\section{Tooling}

\TODO{Eric: high level description of the implementation}

\ERIC{
=> Context VCE, ref to FASE paper\\
=> Current prototype: programatic creation of pNet objects, through the pNet API in VCE. On the middle term, this should be compiled from some language or graphical formalism.\\
=> Open transitions are computed as a direct implementation of the algorithm above. Steps 2a-2b are merged: at each pNet node, after application of rule Tr2, the resulting OT is built from the premisses, without explicitely constructing the deduction tree. \\
=> Variable names are generated for all ``fresh'' and ``clone'' operations, using structured variable names ensuring uniqueness of fresh names, but also reasonable readability of these names for debugging purposes.\\
=> Satisfiability check is done only at toplevel of the deduction tree construction. An alternative would be to submit the satisfiability check to the SMT solver at each level of the tree construction, potentialy reducing the overall number of combinations. But submission to the SMT engine is costly, and more complexity analysis is required before deciding if this would be worthwhile.\\
=> Simplification is not yet implemented. It is not strictly required for the Open Automaton construction, but it will be critical later for predicate comparison in the bisimulation algorithm.}

\TODO{Eric: full result of the tool execution on the running example}

\section{Bisimulation}
\label{section:bisimulation}

	Now we use our symbolic operational semantics to define a notion of
	strong (symbolic) bisimulation. Moreover this equivalence is decidable whenever we 
	have some
	decision procedure on the predicates of the action algebra.

%% , that we prove compatible with the
%% composition of open pNets. \TODO{do we really know this ?}

The equivalence we need is a strong bisimulation between
pNets having exactly the same Holes with the same sorts, but using a
flexible matching 
between open transition, to accommodate comparisons between pNet
expressions with different architectures.
 We name it FH-bisimulation,
 as a short cut for the ``Formal Hypotheses'' manipulated in the
 transitions, but also as a reference to the work of De Simone~\cite{deSimone85},
 that pioneered this idea. Formally:

 \begin{definition}[Strong FH-bisimulation]\label{def-FH-bisim}

\noindent
\begin{minipage}{0.69\linewidth} 	Suppose that
   $A_1 = <L_1,J,\mathcal{S}_1, s^1_0,
   \mathcal{T}_1>$ and $A_2 = <L_2,J,\mathcal{S}_2,s^2_0, \mathcal{T}_2>$
   are open automata where the set
   of holes are equal %\LUDO{added:} 
   and of the same sort.  %\LUDO{changed:} 
      Let 
      \ERIC{$\mathcal{R}=\{(s^1,s^2|\Pred_{s^1,s^2})\}$} be a
      relation over the sets 
      $\mathcal{S}_1$ and $\mathcal{S}_2$ 
      constrained by a predicate.  More precisely, for any pair $(s^1,s^2)$, there is a 
      single
      $(s^1,s^2|\Pred_{s^1,s^2})\in\mathcal{R}$  stating that $s^1$ and $s^2$ are related if $\Pred_{s^1,s^2}$\ 
      is 
      true.
\end{minipage}
\hspace{2mm}
\begin{minipage}{0.30\linewidth}
	\includegraphics[width=\linewidth]{XFIG/Bisim}
\end{minipage}
 Then 
   $\mathcal{R}$ is an FH-bisimulation iff for any  states
   $s^1\in\mathcal{S}_1$ and\\ $s^2\in\mathcal{S}_2$, $(s^1,s^2|\Pred)\in\mathcal{R}$, we 
   have
   the following:

 \begin{itemize}
 \item  For any open transition $OT$ in $\mathcal{T}_1$:
 \begin{mathpar}
     \inferrule*[myfraction=\reddottedrule]
         {\{s^1_i\xrightarrow{a_i}_i s^{1'}_i\}^{i\in I_1}, % , s^1_i,s^{1'}_i\in 
         %States(I_1)
           \{\xrightarrow{b_j}_j\}^{j\in J_1},\Pred_{OT},\Post_{OT}}
         {s^1 \xrightarrow {v} s^{1'}}

\end{mathpar}
 there exist open transitions $OT_x^{x\in X} \subseteq \mathcal{T}_2$:
 \begin{mathpar}
%    \left( fresh \ \overline{a_i}, \overline{b_j}, v_x.\ \
    \inferrule*[myfraction=\reddottedrule]
         {\{s^{2}_i\xrightarrow{a_{i x}}_i s^{2}_{i x}\}^{i\in I_{2 x}},
           \{\xrightarrow{b_{j x}}_j\}^{j\in J_{2x}}, \Pred_{OT_x},\Post_{OT_x}}
         {s^2 \xrightarrow {v_x} s^{2}_x}
%         \right)
\end{mathpar}
 such that  $\forall x, J_1=J_{2x}, (s^{1'},s^{2}_x|\Pred_{s^{1'},s^{2}_x})\in 
 \mathcal{R}$; 
 and  \\
 $\Pred \land \Pred_{OT}\\
\hspace{1cm} \implies\!\!\! \bigvee_{x\in X}
   \left( \forall j. b_j=b_{jx}  \Rightarrow \Pred_{OT_x}
     \land v\!=\!v_x \land  
     \Pred_{s^{1'},s^{2}_x}\subst{\Post_{OT}}\subst{\Post_{OT_x}}\right)$
     %     \symb{Subst}(\Pred_{target_x}, \Post_{OT} o \Post_{OT_x})
     %     \right)$.

 \item  and symmetrically any open transition from $s^2$ in $\mathcal{T}_2$ can be 
      covered by a set of transitions from $t^1$ in $\mathcal{T}_1$.
 \end{itemize}

 \TODO{Eric: il reste des petits bugs genre $s^{2'}_x$ plutot que
   $s^{2}_x$}

%Where $\symb{Subst}(\Pred,\Post)$ is the parallel substitution of all
%assigned variables.
Two pNets are FH-bisimilar if there exist a relation between their associated 
automata that is an FH-bisimulation and their initial states are equivalent.

 \end{definition}
Classically, $\Pred_{target_x}\subst{\Post_{OT}}\subst{\Post_{OT_x}}$
applies in parallel the  
substitutions $\Post_{OT}$ and $\Post_{OT_x}$ (parallelism is crucial
inside each $\Post$ set but $\Post_{OT}$ is independent from
$\Post_{OT_x}$), applying the assignments of the involved rules.

\medskip
\ERIC{If we assume that everything is finite (states and transitions in the
open automata, and the predicates in $\mathcal{R}$, then it is easy to
prove that it is decidable whether a relation is a 
FH-bisimulation provided the logic of the predicates is decidable (see
\cite{henrio:Forte2016}). }

\begin{theorem}[Decidability of FH-bisimulation]
Let $A_1$ and $A_2$ be finite open automata
and $\mathcal{R}$ a relation over their states $\mathcal{S}_1$ and
$\mathcal{S}_2$ constrained by a finite set of predicates. Assume that
the predicates inclusion is decidable over  
the action algebra $\mathcal{A}_P$. Then it is decidable whether the relation 
$\mathcal{R}$ is a FH-bisimulation.

\end{theorem}

\section{Composability}

The main interest of our symbolic approach is to define a method to
prove properties directly on open structures, that will be preserved
by any correct instantiation of the holes. In this section we define a
composition operator for open pNets, and we prove that it preserves
FH-bisimulation. More precisely, one can define two preservation
properties, namely 1) when one hole of a pNet is filled by two bisimilar
other (open) pNets; and 2) when the same hole in two bisimilar pNets are
filled by the same pNet, in other words, composing a pNet with two
bisimilar contexts. The general case will be obtained by
transitivity of the bisimulation relation. We concentrate here on the
second property, that is the most interesting.

\begin{theorem}[FH-Bisimulation is an equivalence]\label{thm-equiv} Suppose $\mathcal{R}$ 
is an FH-bisimualtion. Then $\mathcal{R}$ is an equivalence, that is, $\mathcal{R}$ is 
reflexive, symmetric and transitive.

\end{theorem}

\begin{definition}[pNet composition]
	An open pNet: $\pNet = \mylangle \pNet_i^{i\in I}, S_j^{j\in J}, 
	\overline{\symb{SV}}\myrangle$
 can be (partially) filled by providing  a pNets $\pNet'$ of the
	right sort to fill one of  its holes.	
	Suppose $j_0\in J$:
	%% \[\mylangle \pNet_i^{i\in I},S_j^{j\in J}, \symb{SV}_k^{k\in
	%%   K}\myrangle\left[(\pNet'_i)^{i\in L}\right]= \mylangle \pNet_i^{i\in
	%%   I}\left[(\pNet'_i)^{i\in L}\right]\uplus(\pNet'_i)^{i\in J\cap L},S_j^{j\in 
	%%J\setminus L},
	%% \symb{SV}_k^{k\in K}\myrangle
	%% \]
	\[\pNet\left[\pNet'\right]_{j_0}= \mylangle 
	\pNet_i^{i\in I}\uplus\{j_0\mapsto \pNet'\},S_j^{j\in J\setminus \{j_0\}},
	\overline{\symb{SV}}\myrangle
	\]
\end{definition}

\LUDO{added congruence theorem below}
\begin{theorem}[Congruence]\label{thm-congr-eq}
	Consider an open pNet:
	$\pNet = \mylangle \pNet_i^{i\in I}, S_j^{j\in J}, 
	\overline{\symb{SV}}\myrangle$.
	Let $j_0\in J$ be a hole. Let $Q$ and $Q'$ be two FH-bisimilar pNets such that 
	$\Sort(Q)=\Sort(Q')=S_{j_0}$\footnote{Note that $\Sort(Q)=\Sort(Q')$ is ensured by 
	stronfg bisimilarity.}. Then 
	$\pNet[Q]_{j_0}$ and 
	$\pNet[Q']_{j_0}$ are FH-bisimilar.
\end{theorem}

\begin{theorem}[Context equivalence]\label{thm-ctxt-eq}
	Consider two FH-bisimilar open pNets:
	$\pNet = \mylangle \pNet_i^{i\in I}, S_j^{j\in J}, 
	\overline{\symb{SV}}\myrangle$ and 	$\pNet' = \mylangle {\pNet'}_i^{i\in I}, 
	S_j^{j\in 
	J}, 	\overline{\symb{SV'}}\myrangle$ 
	(recall they must have the same holes to be bisimilar).
	Let $j_0\in J$ be a hole, and $Q$ be a pNet such that $\Sort(Q)=S_{j_0}$. Then $\pNet[Q]_{j_0}$ and 
	$\pNet'[Q]_{j_0}$ are FH-bisimilar.
\end{theorem}

Finally, the previous theorems can be composed to state a general theorem about 
composability and FH-bisimilarity.
\begin{corollary}[Composability]
	Consider two FH-bisimilar pNets with an arbitrary number of holes, when replacing, 
	inside those two original pNets, a subset of the holes by FH-bisimilar pNets, we 
	obtain two FH-bisimilar pNets.
\end{corollary}
%
%In the following we use $\approx$ to denote equality modulo alpha conversion of free 
%variables. In practice, $x\approx y$ can be obtained if: $x=clone(a) \land y=clone(a)$.

\section{Weak bisimulation}

Weak symbolic bisimulation was introduced to relate transition systems
that have indistinguishable behaviour, with respect to some definition
of \emph{invisible actions} that are considered internal to some
subsystem, and consequently cannot be observed, nor used for
synchronisation with their context.
As the notions of invisible action varies in different contexts,
e.g. $tau$ in CCS, and $i$ in Lotos, we will rely here on some set of
\emph{invisible actions} depending on a specific action algebra.

\TODO{we should understand if it make sense to distinguish invisible
  from synchronised actions: in the Forte paper we wrote ``using  as 
\emph{invisible actions} a subset of the
\emph{synchronised actions} defined in Section~\ref{section:pnets}''}.

%% Proving bisimulation properties on our hierarchical broadcast example
%% would take too much space for this paper. Moreover, interesting
%% properties of the HB example would rather be adequate for weak
%% bisimulation, and its large state-space would require some
%% tool-assistance, both for the generation of the open-automata and for
%% their comparison.

Classicaly, we start by defining a derived transition relation based
on sequences of invisible actions:

\begin{definition}[Weak open transitions]
  Let $A = <L,J,\mathcal{S},s_0,\mathcal{T}>$ be an open automaton.
  We define $\mathcal{T^*}$ the \emph{Weak open transition relation}
  with respect to a set of invisible actions $\mathcal{Inv}$ as:

 \begin{mathpar}
 \inferrule*[myfraction=\reddottedrule]
         {\{s_i\xrightarrow{a_i}_i s'_i\}^{i\in I_1}, 
           \{\xrightarrow{b_j}_j\}^{j\in J},\Pred,\Post}
         {s \xRightarrow {v} s'}
         \in \mathcal{T}^*
 \end{mathpar}

 iff there exists a sequence of open transitions
         $OT_x^{x\in X} \subseteq \mathcal{T}$ such that:

 \TODO{Pas si simple !!!}

 \end{definition}

\begin{definition}[Weak FH-bisimulation]\label{def-Weak-bisim}

\noindent
Let $A_1$ and $A_2$ be open automata, and $mathcal{R}$ a relation over
$S_1$ and $S_2$, as in Definition \ref{def-FH-bisim}.

Then 
   $\mathcal{R}$ is an Weak FH-bisimulation iff for any  states
$s^1\in\mathcal{S}_1$ and\\
$s^2\in\mathcal{S}_2$, $(s^1,s^2|\Pred)\in\mathcal{R}$, we 
   have the following:

 \begin{itemize}
 \item  For any open transition $OT$ in $\mathcal{T}_1$:
 \begin{mathpar}
     \inferrule*[myfraction=\reddottedrule]
         {\{s^1_i\xrightarrow{a_i}_i s^{1'}_i\}^{i\in I_1},
           \{\xrightarrow{b_j}_j\}^{j\in J_1},\Pred_{OT},\Post_{OT}}
         {s^1 \xrightarrow {v} s^{1'}}

\end{mathpar}
 there exist open transitions $OT_x^{x\in X} \subseteq \mathcal{T}_2^*$:
 \begin{mathpar}
    \inferrule*[myfraction=\reddottedrule]
         {\{s^{2}_i\xrightarrow{a_{i x}}_i s^{2}_{i x}\}^{i\in I_{2 x}},
           \{\xrightarrow{b_{j x}}_j\}^{j\in J_{2x}}, \Pred_{OT_x},\Post_{OT_x}}
         {s^2 \xRightarrow {v_x} s^{2'}_x}
\end{mathpar}
 such that  $\forall x, J_1=J_{2x}, (s^{1'},s^{2'}_x|\Pred_{s^{1'},s^{2'}_x})\in \mathcal{R}$; 
 and  \\
 $\Pred \land \Pred_{OT}\\
\hspace{1cm} \implies\!\!\! \bigvee_{x\in X}
   \left( \forall j. b_j=b_{jx}  \Rightarrow \Pred_{OT_x}
     \land v\!=\!v_x \land  
     \Pred_{s^{1'},s^{2'}_x}\subst{\Post_{OT}}\subst{\Post_{OT_x}}\right)$

 \item  and symmetrically any open transition from $s^2$ in $\mathcal{T}_2$ can be 
      covered by a set of weak transitions from $t^1$ in $\mathcal{T}_1^*$.
 \end{itemize}

Two pNets are Weak FH-bisimilar if there exist a relation between their associated 
automata that is an Weak FH-bisimulation and their initial sstates are equivalent.

 \end{definition}

\TODO{Here we should include:
  \\  - A definition for Weak FH-Bisimulation
  \\  - A sufficient condition for an open PNet to preserve Weak-FH,
  including: 1) no non-deterministic Tau-choice in the pLTSs, 2) a
  synch vector propagating Tau from a subnet $pNet_i$ mandatory
  everytime $pNet_i$ can do some visible action in the current
  context, 3) if we have a vector <Tau, -, a, - > --> b with ``a'' a visible
  action of some CLOSED subnet, then b=Tau. 
  \\  - the corresponding Congruence theorem
  \\- An example
}

\TODO{Eric: ask Elie, Hubert or others about interesting weak
  equationnal theories for Lotos -- DONE, waiting for answer...}

\TODO{Ludo: ask Simon for weak properties of Bip composition
  ``operators'' -- DONE!}

@@

We are currently extending this work,  looking at  both further properties of 
FH-bisimulation, but also
the relations with existing equivalences on closed systems.
We also plan to apply open pNets to the study of complex composition
operators in a symbolic way, for example in the area of parallel
skeletons, or distributed algorithms.
We have started developping some tool support for computing the
symbolic semantics in term of open-automata. The following steps will
be the development of algorithms and tools for checking 
FH-bisimulations, and interfacing with decision engines for
predicates, typically SMT solvers. Those tools will include
an algorithm that partitions the states and generates the right
conditions (automatically or with user input) for checking
whether two open pNets are bisimilar.

Independently, it is clear that most interesting properties of
such complex systems will not be provable by strong bisimulation.
Next steps will include the investigation of weak
versions of the FH-bisimulation, using the notion of
\emph{synchronised actions} mentionned in the paper.
%We are convinced that this work will contribute to the design of tools for
%verifying open distributed systems.

@@

\newpage
\appendix

\section{Proof of associativity of Enable pNet.}
\label{appendix:assoc}

Using the usual encoding of the Lotos Enable operator (see
Fig. \ref{schema:enable-pnets}), we build now open automata for two  
open pNets \texttt{P>>(Q>>R)} and \texttt{(P>>Q)>>R)}, and 
prove their bisimilarity.
\begin{figure}[t]
\centerline{ % \includegraphics[width=0.5\linewidth]{XFIG/Struct-P-QR}
% \includegraphics[width=0.35\linewidth]{XFIG/Struct-PQ-R}}
  \includegraphics[width=0.60\linewidth]{XFIG/P-QR}
  \hspace{5mm}
 \includegraphics[width=0.27\linewidth]{XFIG/PQ-R}}
  \caption{Structure of the pNets for ``P>>(Q>>R) and (P>>Q)>>R''}
  \label{schema:enable-assoc}
\end{figure}

The two open pNets are drawn in Fig. \ref{schema:enable-assoc}. To
alleviate the drawing, we have only 
  given one instance (C2) of the enable controller pLTSs, and one example of
  the set of synchronisation vectors.

  Below we give the two open automata for these pNets. In the
  open-transitions in these 
    automata, the predicates are very simple, and for readability
    we show them directly in the transitions.
    Remark that while the structure of the automaton states is useful
    for their construction, it is not relevant later, so we have
    dropped this structure when drawing the automaton.

\begin{figure}[h]
  \begin{minipage}{2cm}
  \includegraphics[width=\linewidth]{XFIG/PQR-automaton}
  \end{minipage}
  \hspace{10mm}
  \begin{minipage}{7cm}
  \begin{eqnarray*}
    OT_1 & = \openrule{0 \xrightarrow{l}_{C_3} 0
    	\\\xrightarrow{a1}_P\\ a1\neq\delta(x)}
    {\ostate{00} \xrightarrow{a1} \ostate{00}}\\
    OT_2 & = \openrule{
                       0 \xrightarrow{\delta}_{C_3} 1, ~~
                            0 \xrightarrow{l}_{C_4} 0,  ~~                 
                            \{\xrightarrow{\delta(x1)}_P, ~~
                            \xrightarrow{acc(x1)}_Q\}
                      }
                      {\ostate{00} \xrightarrow{\underline{\delta(x1)}} \ostate{10}}    \\
    OT_3 & = \openrule{0 \xrightarrow{l}_{C_4} 0,
                        \\1 \xrightarrow{r}_{C_3} 1,\\
                        \xrightarrow{a2}_Q,\\a2\neq\delta(x)
                        }
    {\ostate{10} \xrightarrow{a2} \ostate{10}}\\
    OT_4 & = \openrule{
                       0 \xrightarrow{\delta}_{C_4} 1, \qquad
                       1 \xrightarrow{r}_{C_3} 1, \qquad
                       \{\xrightarrow{\delta(x2)}_Q, \qquad 
                       \xrightarrow{acc(x2)}_R\} 
                       }
                      {\ostate{10} \xrightarrow{\underline{\delta(x2)}} \ostate{11}}    \\
    OT_5 & = \openrule{1 \xrightarrow{r}_{C_3} 1, \\1 \xrightarrow{r}_{C_4} 1, 
    	\\\xrightarrow{a3}_R
    	}
    {\ostate{11} \xrightarrow{a3} \ostate{11}}
    \end{eqnarray*}
  \end{minipage}
  \caption{Open Automaton for the term ``P>>(Q>>R)''}
  \label{schema:enable3bis}
\end{figure}

\begin{figure}[t]
  \begin{minipage}{2cm}
  \includegraphics[width=\linewidth]{XFIG/PQR2-automaton}
  \end{minipage}
  \hspace{10mm}
  \begin{minipage}{7cm}
  \begin{eqnarray*}
    OT'_1 & = \openrule
             {
               0 \xrightarrow{l}_{C_3} 0,~~0 \xrightarrow{l}_{C_4} 0,~~
               \{\xrightarrow{b1}_P\},~~ b1\neq\delta(y)
               }
             {\ostate{00} \xrightarrow{b1} \ostate{00}}\\
    OT'_2 & = \openrule
             {
                       0 \xrightarrow{l}_{C_5} 0, ~~
                       0 \xrightarrow{\delta}_{C_6} 1,  ~~
              \{\xrightarrow{\delta(y1)}_P,~~      \xrightarrow{acc(y1)}_Q\}~~ 
                      }
             {\ostate{00} \xrightarrow{\underline{\delta(y1)}} \ostate{01}}\\
    OT'_3 & = \openrule
             {
               0 \xrightarrow{l}_{C_5} 0,\\0 \xrightarrow{r}_{C_6} 0,
               \\ \{\xrightarrow{b2}_Q\},\\b2\neq\delta(y)
               }
    {\ostate{01} \xrightarrow{b2} \ostate{01}}\\
    OT'_4 & = \openrule
             {
                       0 \xrightarrow{\delta}_{C_5} 1, ~~
                       1 \xrightarrow{r}_{C_6} 1, ~~\{\xrightarrow{\delta(y2)}_Q, ~~
                    \xrightarrow{accept(y2)}_R\} 
                                 }
             {\ostate{01} \xrightarrow{\underline{\delta(y2)}} \ostate{11}}    \\
    OT'_5 & = \openrule
             {1 \xrightarrow{r}_{C_5} 1,\\\{\xrightarrow{b3}_R\}
             	}
             {\ostate{11} \xrightarrow{b3} \ostate{11}}
    \end{eqnarray*}
  \end{minipage}
%  \label{schema:enable5}
\end{figure}

In this example, we have no state variables in the pNets leaves, so we
need no auxiliary predicate in the FH-bisimulation.

Let us build a relation
    $R = \{(A1_0,A2_0), (A1_1,A2_1), (A1_2,A2_2)\}$, and prove that
    $R$ is a strong FH-bisimulation.

    Pick e.g. transition $OT_1$ of automaton $A1$, we have
    $A1_0 \xrightarrow{OT_1}_{A1} A1_0$. The matching transition from
    $A2_0$ is $OT'_1$, with 
    $A2_0 \xrightarrow{OT'_1}_{A2} A2_0$, the set of holes involved in
    both transitions is $\{P\}$, and the resulting state is in the
    bisimulation ($A1_0 ~ A2_0$).
    The proof obligation is:

    $$  a1\neq\delta(x) \implies
         a1=b1 \implies b1\neq\delta(y) \land a1=b1 $$

    The proof obligation for $OT_2$ versus $OT'_2$ is still simpler:

    $$ \delta(x_1)=\delta(y_1) \land acc(x_1)=acc(y_1)
       \implies \underline{\delta(x_1)}=\underline{\delta(y_1)} $$

       the 3 other cases, and the 5 in the other direction, are left
       to the reader. All are simple to prove, so $R$ is a
       FH-bisimulation, and the initial states of the 2 pNets are equivalent.
           \qed

       This proof shows how the flattening induced by the open transitions makes the 
       reasoning on hierarchical structures easy, while keeping a structured view of the 
       compared systems.

       \newpage

       \section{Proofs of basic properties on pNets}
       \subsection{Equivalence relation (Theorem~\ref{thm-equiv})}
        \emph{Suppose $\mathcal{R}$ 
       	is an FH-bisimualtion. Then $\mathcal{R}$ is an equivalence, that is, 
       	$\mathcal{R}$ is 
       	reflexive, symmetric and transitive.
       	}
       \begin{proof}

>>>>

%\usepackage{amsfonts}
%\usepackage{fancyhdr}
%\usepackage{cite}
%\usepackage{ifthen}
%\usepackage{amssymb}
%\usepackage{fancyhdr}
%\usepackage{pifont}
\usepackage{stmaryrd}
\usepackage{mathtools,mathpartir}
\usepackage{proof}
%\usepackage{setspace}
%\usepackage{indentfirst}
\usepackage{amsmath,amssymb,amscd,mathrsfs}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}
\usepackage{epsfig,color,subfigure,enumitem}
\newcommand{\TODO}[1]{\textcolor{red}{\textbf{[TODO:#1]}}}
\newcommand{\NOTE}[1]{\textcolor{blue}{\textbf{[NOTE:#1]}}}
\newcommand{\ERIC}[1]{\textcolor{blue}{#1}}
\newcommand{\coloncolon}{{:\hspace{-.2ex}:}}
\makeatletter
\newcommand{\raisemath}[1]{\mathpalette{\raisem@th{#1}}}
\newcommand{\raisem@th}[3]{\raisebox{#1}{$#2#3$}}
\makeatother

@@

\title{Using SMT engine to generate Symbolic Automata\thanks{This work was partially 
funded by the Associated Team FM4CPS
  between INRIA and ECNU, Shanghai}}
\author{ Xudong Qin\inst{3}  \ \ \  Eric Madelaine\inst{1,2}} 
\institute{Univ. of Nice Sophia Antipolis, CNRS, UMR 7271, 06900 Sophia Antipolis, France
	\and INRIA Sophia Antipolis M\'edit\'erann\'ee, BP 93, 06902 Sophia Antipolis, France
\and Shanghai Key Laboratory of Trustworthy Computing, ECNU, China}
\date{}                                           % Activate to display a given date or no date

@@

\begin{abstract}
  We implement the symbolic semantics of open pNets using their
  so-called ``Open Automaton'' behavioural semantics. This involve
  building predicates expressing the synchronisation conditions
  allowing some combination of events in the pNet system. These
  predicates are typically built using first order logic, plus some
  predicates specific of particular action algebras. To reduce the
  complexity of the generated open automata, we use the Z3 SMT engine
  to check satisfiability of the predicates, and prune the state
  space..

@@

\ERIC{imported from FORTE'16, }

@@

\paragraph{Context}
\TODO{
  Summary of previous works on pNets and open pNets => 1/2 page.\\
  Longterm goals:\\
  - open pNets to represent operators and program skeletons (ref to
  \cite{PDP2015})\\
  - compute semantics in term of open transitions\\
  - compute equivalence (bisimulation of open automata)\\
  - model-check properties of open systems
  }
\paragraph{Contribution}

@@

\section{Running Example(s)}

@@

@@

The semantics of open pNets will be defined  as an open automaton. An open
automaton is an automaton where each transition composes transitions of several LTSs with
action of some holes, the transition occurs if some predicates hold, and can involve a 
set of state modifications.

\begin{definition}[Open transitions]
	\label{def:OpenTransitions}
	An \emph{open transition} over a set  $(S_i,s_{0 i}, \rightarrow_i)^{i\in
	I}$ of LTSs, a
	set $J$ of holes with sorts $Sort_j^{j\in J}$, and a set of states $\mathcal{S}$ is a structure of the form:	
	\begin{mathpar}
	\inferrule*[myfraction=\reddottedrule]
	{\{s_i~{\xrightarrow{a_i}}_i ~s_i^{\prime}\}^{i\in I},
		\{\xrightarrow{b_j}_j\}^{j\in J}, \Pred, \Post}
	{s \xrightarrow {v}s'}
	\end{mathpar}
	Where $s, s'\in\mathcal{S}$ and for all
        $i\in I$, $s_i{\xrightarrow{a_i}}_i s_i^{\prime}$ is a transition of the
	LTS $(S_i,s_{0 i}, \rightarrow_i)$, and $\xrightarrow{b_j}_j$
        is a transition of the hole $j$, for any action $b_j$ in the
        sort $Sort_j$. \Pred\ is a predicate
	over the different variables of the
	terms, labels, and states $s_i$, $b_j$, $s$, $v$. \Post\ is a set of equations that 
	hold \emph{after the open transition}, they are represented as a substitution of the 
	form $\{x_k\gets e_k\}^{k\in K}$ 
	where $x_k$ are variables of $s'$, $s'_i$, and $e_k$ are expressions over the other 
	variables of the open transition.
\end{definition}

@@

@@

@@

\subsection{Computing and using open automata}
In this section we present an algorithm to construct the open
automaton representing the behaviour of an open pNet, and we prove that
under reasonable conditions this automaton is finite.

\TODO{Refine the algorithm description}

@@

\section{Implementation}

\TODO{Xudong: Detailed description of the implementation}

\ERIC{
=> Context VCE, ref to FASE paper\\
=> Current prototype: programatic creation of pNet objects, through the pNet API in VCE. On the middle term, this should be compiled from some language or graphical formalism.\\
=> Open transitions are computed as a direct implementation of the algorithm above. Steps 2a-2b are merged: at each pNet node, after application of rule Tr2, the resulting OT is built from the premisses, without explicitely constructing the deduction tree. \\
=> Variable names are generated for all ``fresh'' and ``clone''
operations, using structured variable names ensuring uniqueness of
fresh names, but also reasonable readability of these names for
debugging purposes.\\
=> Assignements\\
=> Satisfiability check is done only at toplevel of the deduction tree construction. An alternative would be to submit the satisfiability check to the SMT solver at each level of the tree construction, potentialy reducing the overall number of combinations. But submission to the SMT engine is costly, and more complexity analysis is required before deciding if this would be worthwhile.\\
=> Simplification is not yet implemented. It is not strictly required for the Open Automaton construction, but it will be critical later for predicate comparison in the bisimulation algorithm.}

\TODO{Xudong: full result of the tool execution on the running example}

@@

@@

% \newpage
% \appendix
